Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-08 surfaces-write-up/surface-write-up.tex
added u-powers

By Proposition ~\ref{prop:cone generation}, in order to bound
the degree of generation of $R_D$, we only need bound
the degrees of extremal rays of an associated cone. We now carry out
this strategy.

@@

\begin{rem}
\label{rem:pm-extremal-rays}
Note for future reference that the extremal rays of the cone $\Sigma$, as
defined in the proof of Lemma ~\ref{lem:proj-generators} correspond to the
elements
\begin{equation}
\label{eqn:epsilon-def-proj}
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}\in R_D.
\end{equation}
\end{rem}

@@

\begin{proof}
We first claim there exist $\beta_1, \ldots, \beta_n \in k[\Sigma]$
such that $\ker \psi$ is generated by
\begin{equation}
\label{eqn:relations-psi-proj}
	(z_i - \beta_i)\prod_{j=0}^n {z_j}^{c_{j}}
\end{equation}

@@

\begin{proof}
%Let $(\bida_i, \bidb_i)$ be the bi-degrees of $f_i$. 
Suppose $g
\in (R_D)_d$ is a monomial.
 Then 
\[
	g = \prod_{i = 1}^n {f_i}^{c_i}
\]

\noindent
for some $c_i \ge - \alpha_i d$ such that $\sum_{i=1}^n c_i \bida_i = 0$
and $\sum_{i=1}^n c_i \bidb_i = 0$. We can view $g$ as an element of the
lattice 
\begin{align*}
	\Sigma = \{ (d, c_1, \ldots, c_n) \in \bz_{\geq 0}^{n+1} : c_i \ge - d
\alpha_i \text{ for all }i \text{ and }\sum_{i=1}^n c_i \bida_i = \sum_{i =
1}^n c_i \bidb_i = 0 \}.	
\end{align*}

@@

\begin{proof}
We first claim that there exist $\beta_1, \ldots, \beta_n\in \bk[z_1, z_2, z_3, z_4]$
 such that $\ker \psi$ has relations of the form
\begin{equation}
\label{eqn:hirz-relations-psi}
	(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

\noindent
lying in degree $d \in \mathbb{N}$
with $c_j \ge -\alpha_j d$ (for all $j$) satisfying $\bida_i + \sum_{j = 1}
^n \bida_j c_j = 0$ and $\bidb_i + \sum_{j=1}^n \bidb_j c_j = 0$.

@@

Consider a relation $(f_i - \beta_i)\prod_{j=1}^n f_j^{c_j}=0$ with corresponding lattice point $\sigma: = (d, c_1, \ldots, c_{i-1}, c_{i}+1, c_{i+1}, \ldots, c_n) \in \Sigma$.
Then we can write $\sigma$ as a sum of $s_j e_j$'s for $j\in
\Te$ and $s_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, defing $d_j := \deg e_j$ (when it exists) and let the $j^{th}$
component of $e_j$ be $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$. Also, let $d_{j,k} := \deg e_{j,k}$ (when it exists) and let the $j^{th}$ component of $e_{j,k}$ be $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and the $k^{th}$
component be $-\alpha_k d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$.

For $(f_i - \beta_i)\prod_{j=1}^n f_j^{c_j}$ to be a relation, it must, in particular, be an element of $R_D$. This implies that $s_i\kappa_i\ge 1$, $\sum_{j\in \Tm} s_{i,j}\kappa_{i,j}' \ge 1$, and $\sum_{j\in \Tp} s_{i,j}\kappa_{j,i}'' \ge 1$ if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$ respectively.

>>>>

By Proposition ~\ref{prop:cone-generation}, in order to bound
the degree of generation of $R_D$, we only need bound
the degrees of extremal rays of an associated cone. We now carry out
this strategy.

@@

\begin{rem}
\label{rem:pm-extremal-rays}
Note for future reference that the extremal rays of the cone $\Sigma$, as
defined in the proof of Lemma ~\ref{lem:proj-generators} correspond to the
elements
\begin{equation}
\label{eqn:epsilon-def-proj}
	\epsilon_i = u^{\ell_i a_i}\prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}\in R_D.
\end{equation}
\end{rem}

@@

\begin{proof}
We first claim there exist $\beta_1, \ldots, \beta_n \in k[\Sigma]$
such that $\ker \psi$ is generated by
\begin{equation}
\label{eqn:relations-psi-proj}
	u^d(z_i - \beta_i)\prod_{j=0}^n {z_j}^{c_{j}}
\end{equation}

@@

\begin{proof}
%Let $(\bida_i, \bidb_i)$ be the bi-degrees of $f_i$. 
Suppose $g
\in (R_D)_d$ is a monomial.
 Then 
\[
	g = u^d \prod_{i = 1}^n {f_i}^{c_i}
\]

\noindent
for some $c_i \ge - \alpha_i d$ for all $i$ such that $\sum_{i=1}^n c_i \bida_i = 0$
and $\sum_{i=1}^n c_i \bidb_i = 0$. We can view $g$ as an element $(d, c_1, \ldots, c_n)$ of the
lattice 
\begin{align*}
	\Sigma = \{ (d', c_1', \ldots, c_n') \in \bz_{\geq 0}^{n+1} : c_i' \ge - d
\alpha_i \text{ for all }i \text{ and }\sum_{i=1}^n c_i' \bida_i = \sum_{i =
1}^n c_i' \bidb_i = 0 \}.	
\end{align*}

@@

\begin{proof}
We first claim that there exist $\beta_1, \ldots, \beta_n\in \bk[z_1, z_2, z_3, z_4]$
 such that $\ker \psi$ has relations of the form
\begin{equation}
\label{eqn:hirz-relations-psi}
	u^d(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

\noindent
lying in some degree $d \in \mathbb{N}$
with $c_j \ge -\alpha_j d$ (for all $j$) satisfying $\bida_i + \sum_{j = 1}
^n \bida_j c_j = 0$ and $\bidb_i + \sum_{j=1}^n \bidb_j c_j = 0$.

@@

Consider a relation $u^d(f_i - \beta_i)\prod_{j=1}^n f_j^{c_j}=0$ with corresponding lattice point $\sigma: = (d, c_1, \ldots, c_{i-1}, c_{i}+1, c_{i+1}, \ldots, c_n) \in \Sigma$.
Then we can write $\sigma$ as a sum of $s_j e_j$'s for $j\in
\Te$ and $s_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, defing $d_j := \deg e_j$ (when it exists) and let the $j^{th}$
component of $e_j$ be $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$. Also, let $d_{j,k} := \deg e_{j,k}$ (when it exists) and let the $j^{th}$ component of $e_{j,k}$ be $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and the $k^{th}$
component be $-\alpha_k d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$.

For $u^d (f_i - \beta_i)\prod_{j=1}^n f_j^{c_j}$ to be a relation, it must, in particular, be an element of $R_D$. This implies that $s_i\kappa_i\ge 1$, $\sum_{j\in \Tm} s_{i,j}\kappa_{i,j}' \ge 1$, and $\sum_{j\in \Tp} s_{i,j}\kappa_{j,i}'' \ge 1$ if $i\in \Te$, $i\in \Tp$, and $i\in \Tm$ respectively.
